Analysis

### 2022-06-03 Logos.Rmd
Trying processing ENCODE-provided PWMs, not use

>>>>

# ENCODE

```{r eval=FALSE}
# Modify input dir
dir_data1 <- file.path(dir_data, "ENCODE_motifs_CTCF")
files <- list.files(dir_data1, pattern = "*.txt", full.names = TRUE)
# Output file
fileNameOut1 <- file.path(dir_data1, "motifs_CTCF.meme")
fileNameOut2 <- file.path(dir_data1, "motifs_CTCF.svg")
# Standard header
write_lines("MEME version 5.4.1 (Tue Mar 1 19:18:48 2022 -0800)", file = fileNameOut1)
write_lines("ALPHABET= ACGT", file = fileNameOut1, append = TRUE)
write_lines("strands: + -", file = fileNameOut1, append = TRUE)
write_lines("Background letter frequencies (from uniform background):", file = fileNameOut1, append = TRUE)
write_lines("A 0.25000 C 0.25000 G 0.25000 T 0.25000 ", file = fileNameOut1, append = TRUE)
# Store number of bases
nsites <- c()
for(i in 1:length(files)) {
  fileNameIn=files[i]
  mtx <- read.table(fileNameIn, sep = " ", fill = TRUE)
  # Make motif name
  motif_name <- strsplit(mtx[1, 1], "\t")[[1]]
  motif_id <- motif_name[1] %>% str_remove(">")
  motif_alt <- motif_name[length(motif_name)]
  # Drop header row
  mtx <- mtx[2:nrow(mtx), ]
  # If non-empty, proceed
  if (nrow(mtx) > 0) {
    # Collect number of bases
    nsites <- c(nsites, nrow(mtx))
    # Motif name
    write_lines(paste0("MOTIF ", paste(motif_id, motif_alt)), file = fileNameOut1, append = TRUE)
    # Header line
    header_line <- paste0("letter-probability matrix: alength= 4 w= ", nrow(mtx), " nsites= 20 E= 0")
    write_lines(header_line, file = fileNameOut1, append = TRUE)
    # Make first numerical column empty
    mtx[, 1] <- ""
    # Write actual data
    write_tsv(mtx, file = fileNameOut1, col_names = FALSE, append = TRUE)
    # Separate from the next motif
    write_lines("", file = fileNameOut1, append = TRUE)
  }
}
print("Number of motifs")
print(length(nsites))
print("Summary of motif widths")
summary(nsites)
# Test if can be imported
motifs_all <- importMatrix(fileNameOut1)
if (!file.exists(fileNameOut2)) {
  dev.off()
  svg(fileNameOut2, width = 10, height = 14)
  motifStack(motifs_all, layout="tree") 
  dev.off()
  }
```

```{r}
# Modify input file
fileNameIn1 <- file.path(dir_data, "motifs_CTCF.txt")
# Output file
fileNameOut1 <- file.path(dir_data1, "motifs_CTCF.meme")
fileNameOut2 <- file.path(dir_data1, "motifs_CTCF.svg")
# Standard header
write_lines("MEME version 5.4.1 (Tue Mar 1 19:18:48 2022 -0800)", file = fileNameOut1)
write_lines("ALPHABET= ACGT", file = fileNameOut1, append = TRUE)
write_lines("strands: + -", file = fileNameOut1, append = TRUE)
write_lines("Background letter frequencies (from uniform background):", file = fileNameOut1, append = TRUE)
write_lines("A 0.25000 C 0.25000 G 0.25000 T 0.25000 ", file = fileNameOut1, append = TRUE)
# Store number of bases
nsites <- c()
for(i in 1:length(files)) {
  file=files[i]
  mtx <- read.table(fileNameIn1, sep = " ", fill = TRUE)
  # Make motif name
  motif_name <- strsplit(mtx[1, 1], "\t")[[1]]
  motif_id <- motif_name[1] %>% str_remove(">")
  motif_alt <- motif_name[length(motif_name)]
  # If non-empty, proceed
  if (nrow(mtx) > 0) {
    # Collect number of bases
    nsites <- c(nsites, nrow(mtx))
    # Motif name
    write_lines(paste0("MOTIF ", motif_name), file = fileNameOut1, append = TRUE)
    # Header line
    header_line <- paste0("letter-probability matrix: alength= 4 w= ", nrow(mtx), " nsites= 20 E= 0")
    write_lines(header_line, file = fileNameOut1, append = TRUE)
    # Make first numerical column empty
    mtx$Pos <- ""
    # Write actual data
    write_tsv(mtx, file = fileNameOut1, col_names = FALSE, append = TRUE)
    # Separate from the next motif
    write_lines("", file = fileNameOut1, append = TRUE)
  }
}
print("Number of motifs")
print(length(nsites))
print("Summary of motif widths")
summary(nsites)
# Test if can be imported
motifs_all <- importMatrix(fileNameOut1)
if (!file.exists(fileNameOut2)) {
  dev.off()
  svg(fileNameOut2, width = 8, height = 2.5)
  motifStack(motifs_all, layout="tree") 
  dev.off()
}
```
